Title

### 2018-07-31 thesis/paper/paper.tex
Added statement non-existence

Negation as a general problem cannot be solved in datalog. One could pre-process every necessary fact to be negated and add this to the ground facts, but this is not the purpose of this work. However, not all constraints require a general negation. Item requires statement, Value requires statement, Inverse and Symmetric are all violated if a statement does not exist on a specific item. This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	Multi-value. The idea is to ask the question if a second statement exists, then evaluating it as above. Both methods will be explained in section \ref{sec_statement_nonexistence}.

@@

The second step is the most complex: Depending on the constrained statement and the requirements, has a violation occurred? For the easiest category of constraints this step does not entail any additional work since the existence of the dependent structure is itself a violation. However, most constraints require at least evaluation of inequality between allowed values and existing values (\ref{sec_inequality}) or the non-existence of certain facts (\ref{sec_statement_nonexistence}). Both processes will be explained in their respective sections.

@@

\section{Statement nonexistence}\label{sec_statement_nonexistence}

>>>>

Negation as a general problem cannot be solved in datalog. One could pre-process every necessary fact to be negated and add this to the ground facts, but this is not the purpose of this work. However, not all constraints require a general negation. Item requires statement, Value requires statement, Inverse and Symmetric are all violated if a statement does not exist on a specific item. This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	Multi-value. The idea is to ask the question if a second statement exists, then evaluating it as above. Both methods will be explained in section \ref{sec_statement_non-existence}.

@@

The second step is the most complex: Depending on the constrained statement and the requirements, has a violation occurred? For the easiest category of constraints this step does not entail any additional work since the existence of the dependent structure is itself a violation. However, most constraints require at least evaluation of inequality between allowed values and existing values (\ref{sec_inequality}) or the non-existence of certain facts (\ref{sec_statement_non-existence}). Both processes will be explained in their respective sections.

@@

\section{Statement non-existence}\label{sec_statement_non-existence}
Multiple statements are violated if specific statements do not exist in relation to constrained statements. It is not possible to simply negate said statements and pre-processing every required statement would defeat the purpose of implementing them in datalog. However, non-existence can be derived utilising the order of statements. Three predicates are introduced: first\([2]\), last\([2]\), and next\([2]\), which denote the first and last statement of an item and the following statement to another using their statement IDs.\\

If qualifier non-existence should be tested the predicates have to be adapted, since they do not have a unique identifier. Their order has to be established using the entire set of statement ID, property and value. For this the predicates first\_qualifier\([3]\) and last\_qualifier\([3]\) taking the statement ID, the property and the value and next\_qualifier\([6]\) with ID, property and value for the current and next qualifier are introduced.\\

Using this pre-processed data the approach is to apply a require-predicate to each statement or qualifier. They will be introduced in the respective constraint sections since they may need to carry different information. For the introduction the predicate require\([1]\) simply denotes that the statement requires something. The qualifier variant works analogous.\\

The pattern consists of two rules, both utilising the same conjunction \(t = \text{tripleEDB}(S, I, P, V) \wedge C_1 .. C_m\) with arbitrary atoms \(C_n\) which is used to test if the triple does not fulfil the requirement. Then rule one
\begin{equation*}
\text{require}(S) \leftarrow \text{first}(S) \wedge t
\end{equation*}
states that the statement requires the condition and
\begin{equation*}
\text{require}(O) \leftarrow \text{next}(S, O) \wedge \text{require}(S) \wedge t
\end{equation*}
propagates this down the statement chain. If the last statement of an item or the last qualifier of a statement has require the violation occurred.\\

In the following constraints only the conjunction \(t\) and adjustments to this pattern will be described. The require-propagation is regarded as given.
